---
title: "Machine Learning Assignment - Week 4 Prediction"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

### Data Source
The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

### Data Load

```{r}
#Load the libraries
suppressMessages(library(caret))
library(rpart)
library(RColorBrewer)
library(RGtk2)
library(rattle)
library(randomForest)

#Load the data
inTrain <- read.csv("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv")
inTest <- read.csv("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv")

#Clean data, remove NA columns. The features containing NA are the variance. Since these values do not add to prediction, they can be removed. We will use this on the training set and assign the same column features to test data set.Remove the first seven features as they are not nummeric or are related to time series.

setdiff(names(inTest[,colSums(is.na(inTest)) == 0]),names(inTrain)) #Difference in column names between train and test sets
colNames <- names(inTest[,colSums(is.na(inTest)) == 0])[8:59]
inTest <- inTest[,c(colNames,"problem_id")]
inTrain <- inTrain[,c(colNames,"classe")]
dim(inTest)
dim(inTrain)
 
```

### Partioning Data Set
```{r}
#We will split the data into 60% training data and 40% test data.We will use the "classe" variable from the train set.
set.seed(12345)
dpTrain <- createDataPartition(inTrain$classe, p=0.6, list=FALSE)
training <- inTrain[dpTrain,]
testing <- inTrain[-dpTrain,]

dim(training)
dim(testing)

```
#### Model Building
For this project, I'll use 2 different model algorithms.
1. Decision trees with CART (rpart)
2. Random forest decision trees (rf)

```{r}
#Build and Predict Decision Tree Model
modFitDT <- rpart(classe ~ ., data = training, method="class")
fancyRpartPlot(modFitDT)

set.seed(12345)
prediction <- predict(modFitDT, testing, type = "class")
confusionMatrix(prediction, testing$classe)

#Build and Predict Random Forest Model
set.seed(12345)
modFitRF <- randomForest(classe ~ ., data = training, ntree = 1000)

prediction <- predict(modFitRF, testing, type = "class")
confusionMatrix(prediction, testing$classe)
plot(modFitRF)
```

### Predicting on Test Data(pml-testing.csv--inTest)
Use the prediction model to predict 20 different test cases

```{r}
#Decision Tree Prediction
predictionDT <- predict(modFitDT, inTest, type = "class")
predictionDT

#Random Forest Prediction
predictionRF <- predict(modFitRF, inTest, type = "class")
predictionRF
```
### Conclusion

Based on an assessment of these 2 model fits and out-of-sample results,random forest model is very accurate, about 99%.